Andres Karjus

### 2022-05-23 index.Rmd
news

<div style="text-align: justify; padding-left:1px"> 
I am a research fellow at the [CUDAN](http://cudan.tlu.ee/) Cultural Data Analytics lab at Tallinn University, working on **language and culture dynamics**, using large corpora, machine learning, and human experiments. Recently I have been also working on, through various collaborations, on art history and creative industries (papers in progress) and advising collaborative projects between involving academic and media industry partners.<br>
I defended my PhD in 2020, in linguistics at the [Centre for Language Evolution](http://www.lel.ed.ac.uk/cle/) of University of Edinburgh, on lexical dynamics and communicative needs in language.
I also occasionally teach **stats and data visualization** for the humanities and social sciences (see [more here](https://andreskarjus.github.io/artofthefigure)). On the right is the (of course slightly outdated) photo, obligatory for all academic home pages, to subtly suggest that in fact I do not spend _all_ my time in the office at all.

@@

* Andres Karjus, Richard A. Blythe, Simon Kirby, Tianyu Wang, Kenny Smith. 2021. Conceptual similarity and communicative need shape colexification". [Cognitive Science](https://doi.org/10.1111/cogs.13035) (open access) | [pdf](https://andreskarjus.github.io/papers/Karjus_2021__Conceptual_Similarity_and_Communicative_Need_Shape_Colexification_An_Experimental_Study__Cognitive_Science.pdf) |  [bib](https://andreskarjus.github.io/bib/karjus_2021_cogsci.bib) | [code and data](https://github.com/andreskarjus/colexification_experiment)
* Andres Karjus, 2020. Competition, selection and communicative need in language change. PhD thesis, University of Edinburgh | [pdf](https://andreskarjus.github.io/thesis/Andres_Karjus_phd_thesis.pdf) | [1-page non-technical summary](https://andreskarjus.github.io/thesis/english_summary.pdf) | [eestikeelne lühikokkuvõte](https://andreskarjus.github.io/thesis/doktoritoo_ylevaade_AK.pdf)
*  Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Communicative need modulates competition in language change | [preprint](https://arxiv.org/abs/2006.09277)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Challenges in detecting evolutionary forces in language change using diachronic corpora. Glossa: a journal of general linguistics, 5(1), p.45. | [open access](http://doi.org/10.5334/gjgl.909) | [code](https://github.com/andreskarjus/wfsim_fit)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2020. Quantifying the dynamics of topical fluctuations in language. Language Dynamics and Change 10(1), 86-125 | [open access](https://doi.org/10.1163/22105832-01001200) | [code](https://github.com/andreskarjus/topical_cultural_advection_model)
* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [journal link](https://doi.org/10.1163/22105832-00802004) | [open postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [journal link](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>

>>>>

<div style="text-align: justify; padding-left:1px"> 
I am a research fellow at the [CUDAN](http://cudan.tlu.ee/) Cultural Data Analytics lab at Tallinn University, working on **language and culture dynamics**, using large corpora, machine learning, and human experiments. Recently I have been also working on, through various collaborations, on art history and creative industries and advising collaborative projects between involving academic and media industry partners.<br>
I defended my PhD in 2020, in linguistics at the [Centre for Language Evolution](http://www.lel.ed.ac.uk/cle/) of University of Edinburgh, on lexical dynamics and communicative need in language.
I also occasionally teach **stats and data visualization** for the humanities and social sciences (see [more here](https://andreskarjus.github.io/artofthefigure)). On the right is the (of course slightly outdated) photo, obligatory for all academic home pages, to subtly suggest that in fact I do not spend _all_ my time in the office at all.

@@

* Andres Karjus, Mar Canet Solà, Tillmann Ohm, Sebastian Ahnert, Maximilian Schich. 2022. Compression ensembles quantify aesthetic complexity and the evolution of visual art | [arxiv preprint](https://arxiv.org/abs/2205.10271) 
* Andres Karjus, Richard A. Blythe, Simon Kirby, Tianyu Wang, Kenny Smith. 2021. Conceptual similarity and communicative need shape colexification". [Cognitive Science](https://doi.org/10.1111/cogs.13035) (open access) | [pdf](https://andreskarjus.github.io/papers/Karjus_2021__Conceptual_Similarity_and_Communicative_Need_Shape_Colexification_An_Experimental_Study__Cognitive_Science.pdf) |  [bib](https://andreskarjus.github.io/bib/karjus_2021_cogsci.bib) | [code and data](https://github.com/andreskarjus/colexification_experiment)
* Andres Karjus, 2020. Competition, selection and communicative need in language change. PhD thesis, University of Edinburgh | [pdf](https://andreskarjus.github.io/thesis/Andres_Karjus_phd_thesis.pdf) | [1-page non-technical summary](https://andreskarjus.github.io/thesis/english_summary.pdf) | [eestikeelne lühikokkuvõte](https://andreskarjus.github.io/thesis/doktoritoo_ylevaade_AK.pdf)
*  Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Communicative need modulates competition in language change | [preprint](https://arxiv.org/abs/2006.09277)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Challenges in detecting evolutionary forces in language change using diachronic corpora. Glossa: a journal of general linguistics, 5(1), p.45. | [open access](http://doi.org/10.5334/gjgl.909) | [code](https://github.com/andreskarjus/wfsim_fit)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2020. Quantifying the dynamics of topical fluctuations in language. Language Dynamics and Change 10(1), 86-125 | [open access](https://doi.org/10.1163/22105832-01001200) | [code](https://github.com/andreskarjus/topical_cultural_advection_model)
* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [journal link](https://doi.org/10.1163/22105832-00802004) | [open postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [journal link](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>
